---
output: 
  word_document:
    reference_docx: template.docx
fontsize: 10pt
params:
  epiweek: 34 # Edit the Week Number for the report here
  year: 2024  # Edit the Year of the report here
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

#Load packages:
pacman::p_load(knitr,        #For Markdown
               tidyverse,    #For data cleaning
               ggplot2,      #For plots and maps
               rio,          #For importing data
               grid,         #For table formatting
               gridExtra,    #For table formatting
               flextable,    #For table formatting
               sf,           #For maps
               glue,         #For text formatting
               officer,      #For rendering formats in Word
               cowplot)      #For combining multiple plots/tables

#Load custom functions: 
source("code/funcs.R")

#Load Data:
maternal_deaths_linelist <- rio::import("data/Maternal deaths.xlsx", which="MD linelist")   #Maternal Deaths: File needs to be replaced for report week
maternal_deaths_summary <- rio::import("data/Maternal deaths.xlsx", which="MD summary")     #Maternal Deaths: File needs to be replaced for report week
weekly_data_by_province <- rio::import("data/weekly_data_by_province.csv")                  #Weekly Data: Contains all weeks through current week
province_shapes <- sf::read_sf("data/shapefiles/zmb_admbnda_adm1_dmmu_20201124.shp")        #Province shapefile for maps
measles_lab <- rio::import("data/measles_lab.csv")                                          #Measles lab results: contains all data through current week
  ## Note: Measles Lab data has dates stored as numeric dates in excel - but with a mix of formats (m/d/y and d/m/y). Manually convert it to csv first then read into R to get the text values instead of Excel date values.
                                          
#Get the start and end dates from the epiweek and year parameters
start_date <- as.Date(paste(params$year, params$epiweek-1, 1, sep = "-"), format = "%Y-%U-%u") #Lag week by 1 to get the date range of the report
end_date <- start_date + 6 

#Basic data tidying
maternal_deaths_linelist_clean <- maternal_deaths_linelist %>%
  janitor::clean_names() %>%
  #Clean date formats
  mutate_at(c("date_of_death"),  ~suppressWarnings(case_when(!is.na(as.Date(as.numeric(.), origin="1899-12-30")) ~ as.Date(as.numeric(.), origin="1899-12-30"),
                                    TRUE ~ parse_date_time(., orders=c("dmy", "mdy"))))) %>%
  mutate_at(c("date_of_notification"),  ~suppressWarnings(case_when(!is.na(as.numeric(.)) ~ as.Date(., "%d/%m/%Y"),
                                    TRUE ~ parse_date_time(., orders=c("dmy", "mdy"))))) %>%
  mutate(age = as.numeric(age)) %>% #Convert age to numeric
  mutate_at(c("primary_cause_of_death", "short_name", "province"), ~stringr::str_to_sentence(tolower(.))) %>% #format text strings for consistency
  mutate(short_name = case_when(short_name == "Non-obstetric complication" ~ "Non-obstetric complications",
                                short_name == "Obstetric hemorrhage" ~ "Obstetric haemorrhage",
                                TRUE ~ short_name)) 
  

maternal_deaths_summary_clean <- maternal_deaths_summary %>%
  janitor::clean_names() %>%
  rename(this_week_total = x2,
         prior_total = x5,
         province = maternal_death_cumulative) %>%
  mutate(cumulative_deaths = rowSums(cbind(this_week_total, prior_total), na.rm = TRUE)) %>% #Combine this week's total with prior cumulative total
  select(province, cumulative_deaths) %>%
  mutate(province = stringr::str_to_title(tolower(province))) %>% #format province name for consistency
  filter(province != "Total") %>% #Remove total row
  arrange(desc(cumulative_deaths))

measles_lab_clean <- measles_lab %>%
  janitor::clean_names() %>%
  mutate_at(
    c("date_of_onset", "date_blood_specimen_collected", "date_specimen_received_in_lab"), 
    ~ suppressWarnings(case_when(
      grepl("^\\d{2}-\\d{2}-\\d{4}$", .) ~ as.Date(., format = "%d-%m-%Y"),  # Convert if in "dd-mm-yyyy" format
      !is.na(as.numeric(.)) ~ as.Date(as.numeric(.), origin = "1899-12-30"),  # Convert numeric to Excel date
      TRUE ~ parse_date_time(., orders = c("dmy", "mdy"))  # Try other common date formats
    ))
  ) %>%
  mutate_at(c("ig_m_results", "rubella_ig_m"), ~case_when(. == 1 ~ "Positive",   #Recode the results using the code shown in the Excel file
                                                          . == 2 ~ "Negative",
                                                          . == 3 ~ "Indeterminate",
                                                          . == 4 ~ "Not done",
                                                          . == 5 ~ "Pending",
                                                          . == 9 ~ "Unknown",
                                                          TRUE ~ "Unknown")) %>%
  mutate_at(c("ig_m_results", "rubella_ig_m"), ~factor(., levels=c("Unknown", "Indeterminate", "Negative", "Not done", "Pending", "Positive"))) %>%
  arrange(province_of_residence, ig_m_results) %>%
  select(-c("x9"))
 
weekly_data_by_province_clean <- weekly_data_by_province %>%
  janitor::clean_names() %>%
  pivot_longer(cols=-c("org_unit_name", "period")) %>%
  mutate(category = case_when(grepl("confirmed", name) ~ "Confirmed",
                             grepl("sent_to_lab", name) ~ "Tested",
                             grepl("suspected", name) ~ "Suspected",
                             TRUE ~ NA_character_)) %>%
mutate(disease = str_remove_all(name, c("_confirmed|_sent_to_lab|_suspected"))) %>%
mutate(disease = case_when(disease == "afp" ~ "AFP",
                           disease %in% c("diarrhoea_non_bloody", "diarrhoea_non_bloodyed") ~ "Non-bloody Diarrhoea",
                           TRUE ~ str_to_title(str_replace_all(disease, "_", " ")))) %>%
  mutate(disease = factor(disease, levels=c("AFP", "Anthrax", "Cholera", "Non-bloody Diarrhoea", "Malaria", "Measles", "Monkeypox", "Plague", "Typhoid Fever"))) %>%
  select(-c("name")) %>%
  pivot_wider(names_from = c("category"), values_from = c("value"), values_fill=0) %>%
  mutate(
    year = as.integer(substr(period, 1, 4)),  # Extract year
    week = as.integer(substr(period, 6, nchar(period))),  # Extract week number
    period2 = lubridate::epiweek(as.Date(ISOdate(year, 1, 1)) + weeks(week - 1))
  )

#Subset all data to weeks 1 through report week, where data allows
maternal_deaths_linelist_clean <- maternal_deaths_linelist_clean %>%
  filter(date_of_notification >= start_date) %>% #Filter to date_of_notification on or after the report start_date
  filter(date_of_notification <= end_date) #Filter to date_of_notification on or before the report end_date

measles_lab_clean <- measles_lab_clean %>%
  filter(date_specimen_received_in_lab <= end_date) #Filter to date_specimen_received_in_lab on or before the report end_date

weekly_data_by_province_clean <- weekly_data_by_province_clean %>%
  filter(year == params$year & week <= params$epiweek) #Filter to weeks on or before report week, within report year

```

```{r header_table, echo=FALSE}

#Format the date for the report header. 
## If start and end dates are in the same month, then display as "Xth - Ynd, Month, Year".
## If start and end dates are in different months, then display as "Xth Month - Ynd Month, Year"
if(format(start_date, "%b") == format(end_date, "%b")){
  report_date_range <- paste0(add_ordinal_suffix(as.numeric(format(start_date, "%d"))), " - ", add_ordinal_suffix(as.numeric(format(end_date, "%d"))), ", ",format(end_date, "%b"),", ",format(end_date, "%Y"))
} else{
  report_date_range <- paste0(add_ordinal_suffix(as.numeric(format(start_date, "%d")))," ", format(start_date, "%b"), " - ", add_ordinal_suffix(as.numeric(format(end_date, "%d")))," ", format(end_date, "%b"),", ",format(end_date, "%Y"))
}

header_data <- data.frame(
  Week = c(paste("Week", params$epiweek)),
  Title = c("Epidemiological Bulletin"),
  DateRange = c(report_date_range)
)

flextable(header_data) %>%
  set_table_properties(layout = "fixed") %>%  # Use fixed layout for precise column control
  width(j = 1, width = 2.5) %>%  # Set left column width
  width(j = 2, width = 2.5) %>%  # Set middle column wider for centering
  width(j = 3, width = 2.5) %>%  # Set right column width
  bg(i = 1, j = 1:3, bg = "#135c2c") %>%  # Dark Green Background for first row
  color(i = 1, j = 1:3, color = "white") %>%  # White Text in first row
  bold(i = 1, j = 1:3) %>%
  fontsize(i = 1, j = 1:3, size = 14) %>%  # Adjust text size
  align(i = 1, j = 1, align = "left") %>%
  align(i = 1, j = 2, align = "center") %>%  # Center Title in its cell
  align(i = 1, j = 3, align = "right") %>%
  valign(i = 1, j = 1:3, valign = "center") %>%  # Ensure vertical centering
  border_remove() %>%
  delete_part(part = "header")

```

```{r sub_header1, echo=FALSE}
sub_heading("Summary", level=1)
```

```{r weekly_notifiable_diseases}
#Totals for this week, by province:
this_week_data_by_province <- weekly_data_by_province_clean %>%
  filter(year == params$year & period2 == params$epiweek) %>%
  group_by(org_unit_name, disease) %>%
  summarise_at(c("Suspected", "Tested", "Confirmed"), ~sum(., na.rm=T)) %>%
  ungroup() %>%
  mutate(province = str_sub(str_remove_all(org_unit_name, " Province"),4)) %>%
  select(province, disease, Suspected) %>%
  filter(Suspected > 0) %>%
  arrange(disease, province)

#Cumulative Totals by disease
cumulative_weekly_data <- weekly_data_by_province_clean %>%
  group_by(disease) %>%
  summarise_at(c("Suspected", "Tested", "Confirmed"), ~sum(., na.rm=T))

#Totals for this week, by disesase:
this_week_data <- weekly_data_by_province_clean %>%
  filter(year == params$year & period2 == params$epiweek) %>%
  group_by(disease) %>%
  summarise_at(c("Suspected", "Tested", "Confirmed"), ~sum(., na.rm=T)) 

#Combine this week and cumulative totals, by disease
weekly_data_summary_table <- this_week_data %>%
  full_join(cumulative_weekly_data, by=c("disease")) %>%
  rename_with(
    ~ c(
      "Disease/Event/Condition",
      paste0("Suspected_Week_", params$epiweek),
      paste0("Tested_Week_", params$epiweek),
      paste0("Confirmed_Week_", params$epiweek),
      "Suspected_Cumulative",
      "Tested_Cumulative",
      "Confirmed_Cumulative"
    ),
    .cols = c(disease, Suspected.x, Tested.x, Confirmed.x, Suspected.y, Tested.y, Confirmed.y)
  ) %>%
  mutate(across(-`Disease/Event/Condition`, ~ format(., big.mark = ",")))

# Construct Sentences for Summary
summary_sentences <- this_week_data_by_province %>%
  group_by(disease) %>%
  summarise(
    total_cases = sum(Suspected),
    locations = paste0(province, "(", Suspected, ")", collapse = ", "),
    .groups = "drop"
  ) %>%
  mutate(
    sentence = glue::glue("{total_cases} suspected cases were reported from {locations} province{ifelse(grepl(',', locations), 's', '')}.")
  ) %>%
  pull(sentence)

summary_sentences <- setNames(as.list(summary_sentences), this_week_data_by_province %>% group_by(disease) %>% summarise(disease=first(disease)) %>% pull(disease))
```

**Immediately Notifiable Diseases and Events**

  - **AFP** : `r summary_sentences["AFP"]`
  - **Anthrax** : `r summary_sentences["Anthrax"]`
  - **Cholera** : `r summary_sentences["Cholera"]`
  - **Measles** : `r summary_sentences["Measles"]`
  
**Other Diseases and Events**

  - **Non-bloody Diarrhoea** : `r summary_sentences["Non-bloody Diarrhoea"]`
  - **Malaria** : `r summary_sentences["Malaria"]`
  - **Typhoid Fever** : `r summary_sentences["Typhoid Fever"]`
  
  
  
```{r sub_header2, echo=FALSE}
sub_heading("Summary Report Priority Diseases, Conditions and Events", level=1)
```

```{r weekly_notifiable_summary_table}
# Store original column names
original_colnames <- names(weekly_data_summary_table)

# Generate second-row header labels
second_row_headers <- original_colnames %>%
  str_replace_all("_Week_\\d+", "") %>%   # Remove "Week_XX"
  str_replace_all("_Cumulative", "")      # Remove "Cumulative"

# Rename columns uniquely to avoid duplication in flextable
colnames(weekly_data_summary_table) <- c(
  "Disease/Event/Condition",
  "Suspected_Week", "Tested_Week", "Confirmed_Week",
  "Suspected_Cum", "Tested_Cum", "Confirmed_Cum"
)

# Create Flextable with Two Header Rows
table1 <- flextable(weekly_data_summary_table) %>%
  set_table_properties(width = 1, layout = "fixed") %>%
  theme_box() %>%
  # First row: Merge "Week X" and "Cumulative Total"
  add_header_row(
    values = c(" ", paste0("Week ", params$epiweek), paste0("Week 1 to ", params$epiweek, ", Cumulative Total")),
    colwidths = c(1, 3, 3)
  ) %>%
  
  # Second row: Rename based on cleaned labels
  set_header_labels(
    `Suspected_Week` = "Suspected",
    `Tested_Week` = "Tested",
    `Confirmed_Week` = "Confirmed",
    `Suspected_Cum` = "Suspected",
    `Tested_Cum` = "Tested",
    `Confirmed_Cum` = "Confirmed"
  ) %>%
  
  # Style the table
  bold(part = "header") %>%
  color(part="header", color="white") %>%
  align(part="all", align = "center") %>%
  align(j=1, align="left", part="body") %>%
  bg(part = "header", bg = "#f79646") %>%
  bold(j = 1, part = "body") %>%
  color(j = 1, color = "white", part = "body") %>%
  bg(j = 1, bg = "#f79646", part = "body") %>%
  border(border = officer::fp_border(color="white"), part="all") %>%
  fontsize(size = 10, part = "all") %>%
  autofit()

table1

```

\pagebreak

```{r sub_header3, echo=FALSE}
sub_heading("Summary of VPD Surveillance Indicators", level=1, top_row_gap=0)
```

```{r sub_header4, echo=FALSE}
sub_heading("Measles Laboratory Test Results by Province", level=2, top_row_gap=0)
```

```{r measles_plot, fig.width=7.5, fig.height=3.2, out.width="100%", out.height="100%"}
#Summarise total samples by result and province
measles_lab_summary <- measles_lab_clean %>%
  group_by(province_of_residence, ig_m_results) %>%
  summarise(total = n()) %>%
  ungroup() 

#Create barchart of total samples by result and province
measles_plot <- ggplot(data=measles_lab_summary) +
  geom_col(aes(x=province_of_residence, y=total, group=ig_m_results, fill=ig_m_results), position="stack") +
  scale_fill_manual(values=c("Positive" = "#ff0000",
                             "Pending" = "#6a477a",
                             "Not done" = "#bebebe",
                             "Negative" = "#0909ff",
                             "Indeterminate" = "#699c68",
                             "Unknown" = "black")) +
  theme_classic() +
  theme(legend.title = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(angle=90, hjust=0.5, vjust=0.5, margin = margin(t = 12)),
        panel.grid.major = element_line(color="grey90", linetype="dashed"),
        plot.margin = margin(25, 10, 25, 10),
        text = element_text(family = "Arial", size = 10))  # Set font for plot)

# Create a Single-Cell TableGrob with Text for side-by-side text and plot
measles_bullet1 <- paste0("The country has recorded a total of ", sum(as.numeric(cumulative_weekly_data$Suspected[cumulative_weekly_data$disease == "Measles"]), na.rm=T), " suspected measles cases in ", params$year,".\n")
measles_bullet2 <- paste0("From the ", sum(as.numeric(measles_lab_summary$total), na.rm=T), " measles specimen that have been tested, ", sum(as.numeric(measles_lab_summary$total[measles_lab_summary$ig_m_results=="Positive"]), na.rm=T), " have been confirmed positive (PR ", round(sum(as.numeric(measles_lab_summary$total[measles_lab_summary$ig_m_results=="Positive"]), na.rm=T)/sum(as.numeric(measles_lab_summary$total), na.rm=T),3)*100,"%)")


# Wrap the text bullets to fit within the allocated space
measles_bullet1_wrapped <- wrap_text_hanging(measles_bullet1, width = 50)
measles_bullet2_wrapped <- wrap_text_hanging(measles_bullet2, width = 50)

text_grob <- tableGrob(
  matrix(c(measles_bullet1_wrapped, measles_bullet2_wrapped), ncol = 1), 
  theme = ttheme_minimal(
  core = list(fg_params = list(fontsize = 10, fontfamily = "Arial", hjust=0, x=0))
))

#Combine the text and plot
cowplot::plot_grid(text_grob, measles_plot, align="h", nrow=1, rel_widths = c(4.5, 4))

```

```{r sub_header5, echo=FALSE}
sub_heading("Maternal Deaths", level=1)
```

```{r maternal_deaths_subheader}
sub_header_data <- data.frame(
  Col1 = c("",  
           paste0("Causes of maternal death (Week ", params$epiweek, ", n=", length(unique(maternal_deaths_linelist_clean$case_id)), ")"),
           ""
  ),
  Col2 = c("",
           paste0("Cumulative distribution of maternal deaths (", params$year, ") by province"),
           "")
)

flextable(sub_header_data) %>%
  set_table_properties(layout = "fixed") %>%  
  width(j = 1:2, width = 3.75) %>%  # Each column gets half of 7.5 inches
  bg(i = 2, j = 1:2, bg = "#f79646") %>%  # Light orange background for second row
  color(i = 2, j = 1:2, color = "white") %>%  # White text
  border(border = officer::fp_border(color="white"), part="all") %>%
  bold(i = 2, j = 1:2) %>%
  fontsize(i = 2, j = 1:2, size = 10) %>%  # Adjust text size
  fontsize(i = 1, j = 1:2, size = 4) %>%  # Adjust the size of the top gap row
  fontsize(i = 3, j = 1:2, size = 4) %>%  # Adjust the size of the bottom gap row
  align(i = 2, j = 1:2, align = "center") %>%
  padding(i = 1, j = 1:2, padding.top = 0, padding.bottom = 0, padding.left = 0, padding.right = 0) %>%  # Adjust padding for top buffer row
  padding(i = 2, j = 1:2, padding.top = 4, padding.bottom = 4, padding.left = 0, padding.right = 0) %>%  # Adjust padding for text row
  padding(i = 3, j = 1:2, padding.top = 0, padding.bottom = 0, padding.left = 0, padding.right = 0) %>%  # Adjust padding for bottom buffer row
  delete_part(part = "header")  # Remove column headers
```

```{r maternal_deaths_plot}
#Summarise maternal deaths this week by cause for barchart, and calculate pct of total deaths
maternal_deaths_plot_data <-  maternal_deaths_linelist_clean %>%
  group_by(short_name) %>%
  summarise(total = n()) %>%
  ungroup() %>%
  mutate(pct = total / sum(total, na.rm=T)) %>%
  arrange(pct)

#Order the maternal deaths causes by pct
maternal_deaths_plot_data <- maternal_deaths_plot_data %>%
  mutate(short_name = factor(short_name, levels=maternal_deaths_plot_data$short_name))

#Create maternal deaths barchart
maternal_death_plot <- ggplot(data = maternal_deaths_plot_data,
       aes(x=short_name, y=total)) +
  geom_col(fill="#65adc2") +
  geom_text(aes(label = scales::percent(pct, accuracy = 1)), 
            position = position_nudge(y = max(maternal_deaths_plot_data$total) * 0.03),  # Nudge labels outside
            hjust = 0,  # Align text left (outside the bars)
            size = 2.5) +  # Adjust text size for better readability
  theme_classic() +
  coord_flip() +
  labs(y="# of maternal deaths") +
  theme(axis.text.y=element_text(size=8),
        axis.title.x=element_text(size=8),
        axis.title.y=element_blank(),
        panel.grid.major = element_line(color="grey90", linetype="dashed")) +
  scale_y_continuous(limits=c(0, max(maternal_deaths_plot_data$total, na.rm=T)+0.8))

```

```{r maternal_deaths_map}
#Combine maternal deaths by province with province shapefile
map_data <- maternal_deaths_summary_clean %>%
  mutate(province = case_when(province == "North Western" ~ "North-Western",
                              TRUE ~ province)) %>%
  full_join(province_shapes, by=c("province" = "ADM1_EN")) %>%
  mutate(label = paste0(province, "(", cumulative_deaths,")")) %>%
  sf::st_as_sf()

#Create maternal deaths map
maternal_death_map <-ggplot(data = map_data) +
  geom_sf(aes(fill = cumulative_deaths)) +
  geom_sf_text(aes(label=label), size=3) +
  scale_fill_gradient(low="white", high="#000eb9")+
  theme_void() +  # Removes all non-data elements
  theme(
    legend.position = c(0.05, 0.95),  # Move legend to top-left (inside)
    legend.justification = c(0, 1),  # Align legend's top-left corner
    legend.background = element_rect(fill = alpha("white", 0), color = NA),  
    legend.key.height = unit(0.25, "in"),  # Reduce legend height
    legend.key.width = unit(0.2, "in"),  # Make legend narrower
    legend.text = element_text(size = 7),  # Reduce text size of legend values
    legend.title = element_text(size = 7),  # Reduce text size of legend title
    plot.margin = margin(0, 0, 0, 0) #Remove all margins
  ) +
  labs(fill="Cumulative Maternal Deaths")

```

```{r maternal_deaths_plot_grid, fig.width=7.5, fig.height=3.5, out.width="100%", out.height="100%"}
#Combine maternal deaths barchart and map into one object
cowplot::plot_grid(
  cowplot::plot_grid(maternal_death_plot, ncol = 1),  # Bar chart
  maternal_death_map,  # Map
  align = "h",
  nrow = 1,
  rel_widths = c(4, 4.5),  # Adjust widths
  scale = c(0.9, 1)  # Scale the bar chart to be slightly shorter
)
```
    
  - The bar chart on the left summarizes the causes of deaths of `r length(unique(maternal_deaths_linelist_clean$case_id))` maternal deaths recorded in week `r params$epiweek`.
  - Hypertensive disorder, Non-obstetric complications, and Obstetric haemorrhage continue to be the leading causes of maternal deaths this year.
  - Cumulatively, in `r params$year`, `r sum(maternal_deaths_summary_clean$cumulative_deaths, na.rm=T)` maternal deaths have been recorded across the country, as depicted on the map.
  - Provinces with darker shades (`r maternal_deaths_summary_clean$province[1]`, `r maternal_deaths_summary_clean$province[2]`, and `r maternal_deaths_summary_clean$province[3]`) indicate those with a higher number of reported maternal deaths.

<!-- Note: the above leading causes of death for the year, in the second bullet, were not available in the provided data. Adjustment may be needed once the full annual maternal death data by disease is available -->
